---
title: "Micromaps"
author: "Dilip Molugu"
date: "December 16, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Setup

```{r}
library(tidyverse)
library(micromapST)
```



```{r}
stateNames <- read.csv(file = 'stateNames.csv',
  header = TRUE,as.is = TRUE)[,1]

toFix <- read.csv(file = "stateNames_aka.csv",
  header = TRUE,as.is = TRUE)
   
head(toFix)
```
## use select to remove varaiables and filter to remove rows
```{r}
mG8Y2011 <- toFix %>% 
  select(-(SigDiff:SigSymbol)) %>%
  filter(Jurisdiction %in% stateNames)
head(mG8Y2011)
```

```{r}
mG8Y2011Fixed <- read.csv(file = "stateNames_aka.csv",
  header = TRUE,as.is = TRUE)

# compare the resulting data frames
all.equal(mG8Y2011, mG8Y2011Fixed)
```


## 1.2 Create a micromap column description data.frame
```{r}
## Run

stateNames <- read.csv(file = 'stateNames.csv',
                       header = TRUE,as.is = TRUE)[,1]
colDesc <- data.frame(
  type = c('id','map','dot','dot', 'dot'),
lab1 = c('','','Observed Deaths','Expected Deaths', 'Potentially excess deaths'),
  col1 = c(NA,NA,'Observed Deaths','Expected Deaths', 'Potentially excess deaths')
  )
```

In the column specificaton above, the micromap is to have
5 types of column.  They are, in left to right order, a map
column, and id column and three dot plot columns.

Each column can have two label lines above it and one
label line below it.  

The lab1 argument is a character string vector with
a character string for each column. The map and id
columns have automatic labels.  We enter two null
string ('', '' ,) as place holders. We want the top
label in the 3rd column to be blank so include another null
string ('',) as a place holder. We want the top labels
in the 4th and 5th columns to be 'Above'.   

The lab2 argument is a character string vector with a
character string for each column.  Again we include
empty character string as place holders for the map
and id columns. This time we label the three last columns.

The lab3 argument ia character string vector with a
character string for each column. The   
dot column bottom labels include the units of measure.

The col1 argument specifies the column values in the data.frame
to locate dots respective columns of the micromap.

A second micromap example specifies dotse as column types. 
Each micromap dotse column requires two columns
from data data.frame, one for mean and one for the standard error. 
plot. Hence the column description will have a col2 line
to specify the "se" data columns.  

There are more panel description arguments that can be included
in the column description file when needed.

## 1.3 Use the micromapST function to produce plot. 

Below the first two arguments are the data data.frame
and the column description data.frame.  

The next two arguments specify column with the
state ids and the type of id. 

The plotNames='full' puts the full statenames in the micromap.
the sortVar = 'MN' uses the variable to sort,
and ascend=FALSE makes the order descending.  
The title argument is a character string of length one or two.  

```{r echo=FALSE, fig.width=7.5, fig.height=10}
#pdf(file= "mG8Y2011Dots.pdf" , width=7.5, height=10)


poppy<-micromapST(mG8Y2011Fixed, colDesc,
           rowNamesCol = 'StateNames', rowNames = 'full',
           plotNames= 'full',
           sortVar = 'Deaths',ascend=TRUE,
           # staggerLab='FALSE'
           axisScale = 'sn',
           staggerLab = FALSE,
           # bordDir=bordGrp,
           title = c('DISEASES',
                     'DATA'))

#dev.off()
```


## 2. Dot plots with 95% confidence interval based on standard error

## 2.1 Change the column description data.frame

The glyph use 'dotse'.  We also specify
the location of the standard errors in the  
data data.frame using the col2 vector.

```{r}
## Run

colDescSe <- data.frame(
type = c('map','id','dotse','dotse','dotse'),
lab1 = c('','','','Above','Above'),
lab2 = c('' ,'','Mean','Basic','Proficient'),
lab3 = c('','','Possible 0-500','Percent','Percent'),
col1 = c(NA,NA,'MN','AB','AP'),
col2 = c(NA,NA,'MN_SE','MN_SE','MN_SE')
  )

t(colDescSe)
```


## 2.2 This time open a pdf file, produce plot, and close the file.  

```{r }
pdf(file = "mG8Y2011DotsSe.pdf" , width = 7.5, height = 10)

micromapST(mG8Y2011, colDescSe,
  rowNamesCol = 1,
  rowNames = "full",
  plotNames = "full",
  sortVar = "MN",ascend = FALSE,
  title = c("NAEP Math Grade 8 in 2011",
          "Average Scale Scores and Achievement Percents"))

dev.off()
```

